The \dealii{} Library, Version 8.5

### 2017-03-03 8.5/paper.tex
List updates to the GMG framework

@@

\hypersetup{
  pdfauthor={
    Wolfgang Bangerth,
    Denis Davydov,
    Timo Heister,
    Matthias Maier,
    Jean-Paul Pelteret
  },
  pdftitle={The deal.II Library, Version 8.5, 2017},
}

@@

\author[4]{Matthias Maier}
\affil[4]{School of Mathematics,
  University of Minnesota,
  127 Vincent Hall, 206 Church Street SE,
  Minneapolis, MN 55455, USA,
  {\texttt{msmaier@umn.edu}}}

\author[2]{Jean-Paul Pelteret}
\affil[2]{Chair of Applied Mechanics, 
  University of Erlangen-Nuremberg, 
  Egerlandstr.\ 5, 
  91058 Erlangen, 
  Germany.
  {\texttt{jean-paul.pelteret@fau.de}}}

@@

The major changes of this release are:
\marginpar{Add one sentence to each of the following.}
\begin{itemize}
\item CellDataStorage class
  TODO: Denis and Jean-Paul

@@

\item Trilinos linear operators?
  TODO: Jean-Paul and Matthias

\item Physics module
  TODO: Denis and Jean-Paul

@@

TODO: Denis and Jean-Paul

@@

\subsection{TODO: Linear operators with Payload and Trilinos}

TODO: Jean-Paul and Matthias

@@

TODO: Jean-Paul

@@

TODO: Mention updated Step-37.

@@

J-P.~Pelteret was supported by the European Research Council (ERC) through the Advanced Grant 289049 MOCOPOLY.

>>>>

\usepackage{pgfplots}
\usepackage{pgfplotstable}

@@

\hypersetup{
  pdfauthor={
    Wolfgang Bangerth,
    Denis Davydov
    Timo Heister
    Martin Kronbichler
  },
  pdftitle={The deal.II Library, Version 8.5, 2017},
}

@@

\author[4]{Martin Kronbichler}
\affil[4]{Institute for Computational Mechanics,
  Technical University of Munich,
  Boltzmannstr.~15, 85748 Garching, Germany.
  {\texttt{kronbichler@lnm.mw.tum.de}}}

@@

The major changes of this release are:
\marginpar{Add one sentence to each of the following.}
\begin{itemize}
\item CellDataStorage class

@@

\item Trilinos linear operators?

\item Physics module

@@

@@

\subsection{Linear operators with Trilinos?}

@@

\subsection{Scalability of geometric multigrid framework}

For the new release, the geometric multigrid in \dealii{} have been thoroughly
overhauled regarding their scalability on large-scale parallel computers. To
this end, the geometric multigrid algorithm based on the fast matrix-free
kernels from \cite{KronbichlerKormann2012} have been benchmarked up to
147,456 cores. Several scalability bottlenecks have been removed, including
unnecessary inner products inside the Chebyshev smoother and
$\mathcal O(n_\text{levels})$ global communication steps during the
restriction process rather than only the single global communication step that
is necessary when going to the coarest grid. New matrix-free transfer
implementations called \texttt{MGTransferMatrixFree} were devised that replace
the matrix-based \texttt{MGTransferPrebuilt}. Besides better scalability than
the Trilinos Epetra matrices underlying the latter, the matrix-free transfer
is also a much better for high-order elements with complexity per degree of
freedom of $\mathcal O(d p)$ in the polynomial degree $p$ in $d$ dimensions
rather than $\mathcal O(p^d)$ for the matrix-based approach.

\begin{figure}
\pgfplotstableread{
nprocs     fem256k    fem2m     fem16m    fem128m  fem1g    fem8g
16         0.640934   4.97684   nan       nan      nan      nan
32         0.325741   2.51771   nan       nan      nan      nan
64         0.1645     1.2823    nan       nan      nan      nan
128        0.090898   0.658832  5.00366   nan      nan      nan
256        0.059922   0.339999  2.56216   nan      nan      nan
512        0.0455449  0.176482  1.29986   nan      nan      nan
1024       0.0368049  0.099691  0.67364   6.48155  nan      nan
2048       0.0348921  0.069573  0.356066  2.59601  nan      nan
4096       0.0367949  0.056833  0.19293   1.3251   nan      nan
8192       0.033958   0.045350  0.110485  0.790214 5.50143  nan
16384      0.0379629  0.049351  0.099904  0.424692 2.81479  nan
32768      0.0461671  0.051276  0.077546  0.229114 1.53091  nan
65536      0.0466189  0.058941  0.075194  0.127353 0.819    6.2909
147456     nan        nan       nan       0.087949 0.43213  2.90856
}\scalinglarge
\pgfplotstableread{
nprocs    newdg256k    newdg2m   newdg16m olddg256k  olddg2m   olddg16m
28        1.4398       12.2014   nan      1.458725   12.25674  nan
56        0.6987        6.2015   nan      0.721993   6.266567  nan
112       0.3352        3.1782   nan      0.353412   3.263585  nan
224       0.1888        1.5687   12.8235  0.183643   1.611967  13.03189
448       0.0853        0.7686   6.5649   0.105388   0.805367  6.700762
896       0.0478        0.3537   3.3685   0.063270   0.383875  3.446427
1792      0.0317        0.1717   1.7061   0.046065   0.199738  1.762310
3584      0.0235        0.1079   0.8580   0.042032   0.123113  0.902782
7168      0.0207        0.0668   0.4328   0.037678   0.091822  0.462007
14336     0.0183699     0.045095 0.23276  0.038804   0.071644  0.288911
}\scalingHSW
\centering
\definecolor{gnuplot@green}{RGB}{0,158,115}
\begin{tikzpicture}
    \begin{loglogaxis}[
      title style={at={(0.5,0.965)},anchor=north,draw=black,fill=white,font=\scriptsize\bf},
      title={strong and weak scaling, continuous $\mathcal Q_3$ elements},
      width=0.53\textwidth,
      height=0.5\textwidth,
      xlabel={Number of cores},
      ylabel={Solver time [s]},
      x label style={at={(0.5,0.02)}},
      y label style={at={(0.05,0.5)}},
      xtick={32,128,512,2048,8192,32768,147456},
      xticklabels={32,128,512,2048,8192,32k,147k},
      tick label style={font=\scriptsize},
      label style={font=\scriptsize},
      legend style={font=\scriptsize},
      legend pos=south west,
      ymin=5e-3, ymax=15,
      xmin=8, xmax=147456,
      grid
      ]
      \addplot table[x={nprocs}, y={fem8g}] {\scalinglarge};
      \addlegendentry{8B cells};
      \addplot table[x={nprocs}, y={fem1g}] {\scalinglarge};
      \addlegendentry{1B cells};
      \addplot table[x={nprocs}, y={fem128m}] {\scalinglarge};
      \addlegendentry{128M cells};
      \addplot table[x={nprocs}, y={fem16m}] {\scalinglarge};
      \addlegendentry{16M cells};
      \addplot[gnuplot@green,mark=diamond*,mark options={fill=gnuplot@green!40}] table[x={nprocs}, y={fem2m}] {\scalinglarge};
      \addlegendentry{2M cells};
      \addplot[dashed,black] coordinates {
        (8,10)
        (147456,5/9168)
      };
      \addlegendentry{linear scaling};
      \addplot[dashed,black] coordinates {
        (16,8*5)
        (147456,8*5/9168)
      };
      \addplot[dashed,black] coordinates {
        (16,64*5)
        (147456,64*5/9168)
      };
      \addplot[dashed,black] coordinates {
        (16,5*512)
        (147456,5*512/9168)
      };
      \addplot[dashed,black] coordinates {
        (16,5*4096)
        (147456,5*4096/9168)
      };
    \end{loglogaxis}
  \end{tikzpicture}
  \begin{tikzpicture}
    \begin{loglogaxis}[
      title style={at={(1,0.965)},anchor=north east,draw=black,fill=white,font=\scriptsize\bf},
      title={$256^3$ mesh, discontinuous $\mathcal Q_3$ elements},
      width=0.48\textwidth,
      height=0.5\textwidth,
      xlabel={Number of cores},
      x label style={at={(0.5,0.02)}},
      xtick={56,224,896,3584,14336},
      xticklabels={56,224,896,3584,14336},
      tick label style={font=\scriptsize},
      label style={font=\scriptsize},
      legend style={font=\scriptsize},
      legend pos=south west,
      xmin=28, xmax=14336,
      ymin=5e-3, ymax=15,
      grid
      ]
      \addplot[blue,mark=*,densely dashed] table[x={nprocs}, y={olddg16m}] {\scalingHSW};
      \addlegendentry{old, 16M cells};
      \addplot[blue,mark=o] table[x={nprocs}, y={newdg16m}] {\scalingHSW};
      \addlegendentry{new, 16M cells};
      \addplot[red,mark=square*,densely dashed] table[x={nprocs}, y={olddg2m}] {\scalingHSW};
      \addlegendentry{old, 2M cells};
      \addplot[red,mark=square] table[x={nprocs}, y={newdg2m}] {\scalingHSW};
      \addlegendentry{new, 2M cells};
      \addplot[gnuplot@green,mark=diamond*,densely dashed] table[x={nprocs}, y={olddg256k}] {\scalingHSW};
      \addlegendentry{old, 256k cells};
      \addplot[gnuplot@green,mark=diamond] table[x={nprocs}, y={newdg256k}] {\scalingHSW};
      \addlegendentry{new, 256k cells};
    \end{loglogaxis}
  \end{tikzpicture}
  \caption{Scaling of multigrid algorithms on SuperMUC.}
\label{fig:scaling_mg}
\end{figure}

The scalability of the improved geometric multigrid framework is shown in
Fig.~\ref{fig:scaling_mg}, including a combined strong and weak scalability
plot in the left panel using continuous $\mathcal Q_3$ elements on 57 million
to 232 billion degrees of freedom for the Laplacian. Along each line, the same
problem size is solved with an increasing number of cores, whereas different
lines always start out at 3.5 million degrees of freedom per core. Almost
ideal scalability down to approximately 0.1 seconds can be observed also on
147k cores. The right panel of Fig.~\ref{fig:scaling_mg} shows the effect of
the aforementioned algorithmic improvements on a setup with discontinuous DG
elements, clearly improving the latency of the multigrid V-cycle. The improved
algorithms are shown in the updated step-37 tutorial program.

@@

The framework was also included in the updated step-37 tutorial program.

@@

M.~Kronbichler was partially supported by the German Research Foundation (DFG)
under the project ``High-order discontinuous Galerkin for the exa-scale''
(ExaDG) within the priority program ``Software for Exascale Computing''
(SPPEXA), grant agreement no.~KR4661/2-1, the Bayerisches Kompetenznetzwerk
f\"ur Technisch-Wissenschaftliches Hoch- und H\"ochstleistungsrechnen
(KONWIHR), and the Gauss Centre for Supercomputing e.V.~by providing computing
time on the GCS Supercomputer SuperMUC at Leibniz Supercomputing Centre (LRZ)
through project id pr83te.
